---
params:
title: 'Resources benchmarks'
output:
  # distill::distill_article:
  #   theme: hpstr
  #   toc: yes
  #   code_folding: true
  #   keep_md: true
  html_document:
    toc: true
    code_folding: hide
    toc_float:
      collapsed: true
    theme: flatly
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
.packages <- c("tidyverse")
.bioconductor_packages <- c()

# Install CRAN packages (if not already installed)
.inst <- .packages %in% installed.packages()
if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst], dependencies = TRUE, quiet = TRUE)

# Install Bioconductor packages (if not already installed)
.inst <- .bioconductor_packages %in% installed.packages()
if(length(.bioconductor_packages[!.inst]) > 0) BiocManager::install(.bioconductor_packages[!.inst], dependencies = TRUE, ask = FALSE)

# Load packages into session 
lapply(.bioconductor_packages, require, character.only=TRUE)
lapply(.packages, require, character.only=TRUE)

# set ggplot theme
theme_set(theme_minimal())
```

```{r}
# read in benchmarks file
d = read_tsv("./../../results/stats/benchmarks.tsv", show_col_types = F)
# add column for decimal hours
d$h = d$s / 3600
# same for minutes
d$m = d$s / 60
# convert MB to GB in max_rss
d$max_rss = d$max_rss / 1024
```

# Meaning of all recorded measurements

recorded by Snakemake using ``psutil``

| colname | type (unit) | description |
|-----------|-------------|-----------------------------------------------------------|
| s | float (seconds) | Running time in seconds |
| h:m:s | string (-) | Running time in hour, minutes, seconds format |
| max_rss | float (MB) | Maximum "Resident Set Size”, this is the non-swapped physical memory a process has used. |
| max_vms | float (MB) | Maximum “Virtual Memory Size”, this is the total amount of virtual memory used by the process |
| max_uss | float (MB) | “Unique Set Size”, this is the memory which is unique to a process and which would be freed if the process was terminated right now. |
| max_pss | float (MB) | “Proportional Set Size”, is the amount of memory shared with other processes, accounted in a way that the amount is divided evenly between the processes that share it (Linux only) |
| io_in | float (MB) | the number of MB read (cumulative). |
| io_out | float (MB) | the number of MB written (cumulative). |
| mean_load | float (-) | CPU usage over time, divided by the total running time (first row) |
| cpu_time | float (-) | CPU time summed for user and system |


# Time distribution

```{r}
# boxplots of time per rule
d %>% 
  ggplot(aes(x = rule, y = m, fill = rule)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position="none") +
  labs(x = "Rule",
       y = "Time (min)")
```


# RAM usage distribution

Using max_rss (Maximum Resident Set Size) as measurement for RAM/memory.  

```{r}
# boxplots of max_rss per rule with log10 scale
d %>% 
  ggplot(aes(x = rule, y = max_rss, fill = rule)) +
  geom_boxplot() +
  scale_y_log10(
    limits = c(1e-2, 1e2),
    breaks = c(1e-2, 1e-1, 1, 10, 100),
    labels = c("10MB", "100MB", "1GB", "10GB", "100GB"),
  ) +
  theme(legend.position="none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Rule",
       y = "max_rss")
```


# Recommended values for resources

by taking the maximum +10% and rounding up

```{r}
d %>%
  group_by(rule) %>%
  summarise(
    memory_GB = (1.05*max(max_rss, na.rm = T)) %>% round(),
    time_minutes = (1.05*max(m, na.rm = T)) %>% round()
  ) %>%
  knitr::kable()
```
